Theory about Normal Basis on $F(2^k)$

### 2016-12-11 dissertation/recha.tex
dissertation v3:milestone

\chapter{Word-level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, especially for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew or power
density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating}.
% some major refactoring and attachments are made on original designs. 
These modifications may introduce malfunctions to the original logic and cause problems. 
Based on traditional localized simulation or formal verification method (e.g. equivalence checking), 
designers are reluctant to make aggressive optimization
since the malfunctions are considered as ``faults" in this circuit.
However, if the circuit behavior is carefully investigated, it may become evident that 
those ``faults" will never be activated/excitated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those ``faults" as 
{\bf spurious faults} (false negatives), since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constraints upon the machine to make it start from designated initial states, and 
take specific legal inputs, a set of reachable states can be derived. 
As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
spurious faults by checking if they belong to the reachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One among them is 
state space traversal. Conventionally the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
Before introducing the details of our approach, there are a couple of questions to ask: why is it
necessary to execute word-level FSM traversal? Why does algebraic geometry make this happen?
The answers can be found in this section, as statement of the motivation of our research.

\subsection{FSM Traversal Algorithms}
Sequential circuits are modeled as FSMs, which can be implemented as graph. Thus a 
graph-traversal based algorithm is created to analysis the reachable states \cite{coudert2003unified}.
A traversal algorithm using concept ``implicit state enumeration" is proposed \cite{cho1993redundancy}.
Concretely, the algorithm is written as follows:

\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Sequential circuit with given number of registers}
 \KwOut{Scan registers listed in decreasing order of their non-controllability}

  $from^0 = reached = S^0$\;
  $i = 0$\;
  \While{TRUE}
  {
  	$i++$\;
	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
	$new^i = to^i \cdot \overline{reached}$\;
	\For{each state variable $r_j$}
	{
		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
	}
	\If{$new^i == 0$}
	{
		break\;
	}
	$reached = reached + new^i$\;
	$from^i = new^i$\;
  }\
\Return{$reached$}
\caption {FSM Traversal using Implicit State Enumeration\cite{KallaPartialScan}}
\label{alg:SIMPSON}
\end{algorithm}
\DecMargin{1em}

@@

Since a typical sequential circuit usually contains a combinational logic
component, the traversal algorithm analyzes the combinational logic and derives the transition 
function for one-step reachability within current time-frame, and extends the result to complete 
execution paths through unrolling. If each state encoding (i.e. exact values in the selected registers) is
explicitly analyzed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the BFS traversal algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjunction of several 
Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.

@@

\subsection{Word-level Data Flow on Modern Datapath Designs}
The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
modern digital IC designs such as powerful central processing units (CPUs) and high bandwidth memory (HBM).
Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
is described by the designer as word-level specifications. Therefore, it will be straightforward 
and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
techniques will greatly reduce the state space and make verification more efficient.

@@

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\Indm
 \KwIn{Multiplicand $A,B$}
 \KwOut{Product $C$}
\Indp

@@

We can see that when checking a single part of the whole structure, the number of clauses 
needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
carry-in will become longer and longer, the final conjunction of all clauses will contain 
$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables
\begin{equation}
\label{eqn:wordspec}
S=P_{1/2}+B
\end{equation}

Since the abstraction to word-level reduces symbolic storage and execution cost, the complexity 
to traverse the state space is greatly reduced.
\end{Example}

On the other hand, when implementation details of the datapath are not available, it is not convenient 
any more for users of conventional model checker. The reason is that the user has to write all clauses 
for the implementation, which contains cross-literals, e.g. $s_2$ may associate with $a_1, p_0$, etc. If the user 
is not familiar with the implementation of this adder, those cross-literals will bring
confusions. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
the verification tool will be very user-friendly and straightforward even if the implementation details
are in a black box.

\subsection{On the Existence of Word-level Abstraction}
When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
to word-level first. This conversion is usually completed by abstraction techniques.

An old but universally effective abstraction method is {\it Lagrange's interpolation}. Instead of using 
it in real algebra, it can also be extended to Galois field. Here we use an example to illustrate the 
conversion in Galois field using Lagrange's interpolation.

@@

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
Now our objective is to abstract a function over GF $\F_{2^3}$ about word-level variables, i.e. $Z = \Func(A)$.
Recall the definition of Lagrange's interpolation:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

The Lagrange's interpolation theorem also proves the existence of a word-level abstraction for a bit-level netlist.
In practice, Lagrange's interpolation is not scalable. 
The reason is that it needs entire function (state space), but usually we only have the circuit representation
of the FSM. Considering this fact, a symbolic method is needed.
Our approach in this section uses abstraction based on 
Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in Section \ref{sec:abstraction}.

\subsection{Significance of Developing Word-level FSM Traversal}
At the end of this section, we summarize each subsection and conclude the motivation of our research.
First, we start from the basic FSM traversal algorithms and exhibit their
difficulties to overcome. Secondly, we state the observation that word-level description is 
increasingly important and common in characterizing the data flow on modern large-size datapaths. 
Last but not least, we give instances that some prerequisites for supporting word-level techniques 
are already available.

To conclude, first we state the importance of performing FSM traversal in sequential circuit reachability analysis.
To overcome the vast cost brought by searching in a large space, we propose to use word-level 
polynomials to represent the states and transition relations. As a result, states are categorized into 
sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
and multiple transition relations are therefore merged together. All of these efforts reduces the space cost of 
state space, meanwhile lower the time complexity to traverse such a state space. By using Lagrange's interpolation,
we prove the feasibility of word-level techniques on general circuits' verification.
Thus we cover both the necessity and sufficiency of developing such 
a word-level traversal technique in our work.
They are also answers to the questions at the beginning of this section.

@@

\subsection{FSM model for sequential circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, i.e. the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$
\begin{figure}[H]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{newfig/seqmodel.eps}
% \vspace{-0.2in}
\caption{FSM models of sequential circuits}
%\end{minipage}
\label{fig:seqmodel}}
\end{figure}
In some cases, arithmetic computations are implemented as Moore machines where input operands
are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
We can simplify them to the model in Figure \ref{fig:seqmodel}(b).

@@

Here, $\exists x (f)$ represents the {\it existential quantification
  of $f$ w.r.t. variable $x$}. In Boolean logic, this operator is implemented as
  $$\exists x (f) = f_x\lor f_{\overline{x}}$$

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\Indm
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
\Indp

@@

\subsection{FSM Traversal at word-level over $\Fkk$} 
The state transition graph (STG) shown in Figure \ref{fig:fsm}(a) uses a
2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$. Here, we take 
$P(X) = X^2+X+1$ as the irreducible polynomial to construct
$\mathbb{F}_4$, and $P(\alpha) = 0$ so that $\alpha^2 + \alpha + 1 =
0$.

{\it Initial state:} In Line 1 of Algorithm \ref{alg:BFS}, the initial state is
specified by means of a corresponding polynomial $f = \mathcal{F}(S) =
S - 1 - \alpha$. Notice that if we consider the ideal generated by the
initial state polynomial, $I = \langle f\rangle$, then its variety
$V(I) = 1+\alpha$ corresponds to the state encoding $S_3 = \{11\} =
1+\alpha$, where a polynomial in word-level variable $S$ encodes the
initial state. 
% with only one generator $f$, its variety
% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
% valid value $S_3$ can take.

@@

Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
i.e. the variety of vanishing ideal contains
 all possible valuations of variables, so it constitutes the {\bf
   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
 the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
 computed as:

@@

@@

\subsection{Word-level FSM Traversal Example}

\begin{Example}
\label{ex:SMPO}
We apply Algorithm \ref{alg:univa} to the example shown in
Figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F_4$. Polynomially, it is
written as $from^0 = S - 0$. In the first iteration, we compose
an ideal $J$ with

@@

$V(\langle T^3 + 1\rangle) = \{1, \alpha, \alpha+1\}$ denoting the states
$\{01,10,11\}$. Then the newly reached state set in this iteration is 
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle
T^2+(\alpha+1)T+\alpha \rangle$$ We add these states 
to formerly reached states 
\begin{align*}
reach &= \langle T\rangle \cdot \langle T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T^3+(\alpha+1)T^2+\alpha T\rangle
\end{align*}
 i.e. states $\{00,01,10\}$. We update the present states
for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$

>>>>

\chapter{Word-Level Traversal of Finite State Machines using Algebraic Geometry}
\label{ch:reacha}
Reachability analysis is a basic component of sequential circuit verification, especially for 
formal equivalence checking and model checking techniques. Concretely, in modern synthesis 
tools, in order to improve various performance indicators such as latency, clock skew or power
density, sequential optimization techniques such as retiming \cite{retiming}, scan logic \cite{scan},
sequential power optimization \cite{mathur2009power} and clock-gating techniques \cite{clockgating} are applied.
% some major refactoring and attachments are made on original designs. 
These modifications may introduce, bugs, errors or malfunctions to the original logic and cause problems. 
Based on traditional localized simulation or formal verification method ({\it e.g.} equivalence checking), 
designers are reluctant to make aggressive optimization
since the malfunctions are considered as ``faults" in the circuits.
However, if the circuit behavior is carefully investigated, it may become evident that 
those ``faults" may never be activated during a restricted execution starting 
from legal initial states and with legal inputs. Thus we will call those ``faults" as 
{\it spurious faults} (false negatives), since they will not affect the circuit's normal behavior.

Almost all practical sequential logic components can be modeled as finite state machines (FSMs). 
If we apply constraints upon the machine to make it start from designated initial states, and 
take specific legal inputs, a set of reachable states can be derived. 
As long as the ``faults" can be modeled as ``bad states", we can judge whether they are 
spurious faults by checking if they belong to the unreachable states. From the spurious fault validation 
perspective, reachability analysis is a must when developing full set of sequential circuit verification
techniques.

There are quite a few methods to perform reachability checking on FSMs. One among them is 
state space traversal. Conventionally, the algorithm is based on bit-level techniques such as
binary decision diagrams (BDDs) and Boolean logic. We propose a new traversal algorithm on word-level,
which brings critical advantages. In this chapter the approach will be described and discussed in depth, 
with examples and experiments showing its feasibility when applied on general circuit benchmarks.

\section{Motivation}
Before introducing the details of our approach, there are a few questions to ask: what is 
the benefit of execute FSM traversal at word-level? How can algebraic geometry make this happen?
The answers can be found in this section, as a statement of the motivation of our research.

\subsection{FSM Traversal Algorithms}
Sequential circuits are modeled as FSMs, which can be implemented as graphs. Thus a 
graph-traversal based algorithm is created to analysis the reachable states \cite{coudert2003unified}.
A traversal algorithm using the concept of implicit state enumeration is proposed \cite{cho1993redundancy}.
Concretely, the algorithm is given in Algorithm \ref{alg:BFS}:

% \begin{algorithm}[H]
% \SetAlgoNoLine
% \LinesNumbered
%  \KwIn{Sequential circuit with given number of registers}
%  \KwOut{Scan registers listed in decreasing order of their non-controllability}
% 
%   $from^0 = reached = S^0$\;
%   $i = 0$\;
%   \While{TRUE}
%   {
%   	$i++$\;
% 	$to^i = $IMAGE$(\Delta,from^{i-1})$\;
% 	$new^i = to^i \cdot \overline{reached}$\;
% 	\For{each state variable $r_j$}
% 	{
% 		record\_if\_transitions\_present\_or\_missing$(r_j,new^i)$\;
% 		compute\_degree\_of\_unsettability$(r_j,new^i)$\;
% 	}
% 	\If{$new^i == 0$}
% 	{
% 		break\;
% 	}
% 	$reached = reached + new^i$\;
% 	$from^i = new^i$\;
%   }\
% \Return{$reached$}
% \caption {FSM Traversal using Implicit State Enumeration\cite{KallaPartialScan}}
% \label{alg:SIMPSON}
% \end{algorithm}
% \DecMargin{1em}

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
\Indp

@@

Since a typical sequential circuit usually contains a combinational logic
component, the traversal algorithm analyzes the combinational logic and derives the transition 
function for one-step reachability within current time-frame, and extends the result to complete 
execution paths through unrolling. If each state encoding ({\it i.e.} exact values in the selected registers) is
explicitly analyzed and counted during the unrolling 
procedure, this unrolling is called {\bf explicit unrolling}. In the BFS traversal algorithm, the states cannot be 
directly read in the execution; instead, they are implicitly represented using a conjunction of several 
Boolean formulas. Such techniques differs from explicit unrolling are called {\bf implicit unrolling}.

@@

\subsection{Word-Level Data Flow on Modern Datapath Designs}
The level of integration of modern digital circuit designs is very high. For example, processor A10 designed by Apple
integrates 3.3 billion transistors on a $125~mm^2$ chip \cite{AppleA10}. Such a high density makes the silicon 
implementation of large datapaths possible. In recent decades, 64-bit or even larger datapaths frequently appear in 
modern digital IC designs such as powerful central processing units (CPUs) and high bandwidth memory (HBM).
Meanwhile, with the development of electronic design automation (EDA) tools, data flow 
is described by the designer as word-level specifications. Therefore, it will be straightforward 
and beneficial for users if formal verification tools can work on word-level. Moreover, adopting word-level 
techniques will greatly reduce the state space and make verification more efficient.

@@

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{Multiplicand $A,B$}
 \KwOut{Product $C$}
\Indp

@@

We can see that when checking a single part of the whole structure, the number of clauses 
needed will increase to $k+1$ when using $k$-bit datapath. Considering the formula representing 
carry-in will become longer and longer, the final conjunction of all clauses will contain 
$O(2^k)$ Boolean operators. If by some means we can write the specification with only 3 variables:
\begin{equation}
\label{eqn:wordspec}
S=P_{1/2}+B
\end{equation}

The abstraction to word-level will reduce symbolic storage and execution cost; the complexity 
to traverse the state space will be greatly reduced.
\end{Example}

On the other hand, when implementation details of the datapath are not available, it is not convenient 
any more for users of conventional model checker. The reason is that the user has to write all clauses 
for the implementation, which contains cross-literals, {\it e.g.} $s_2$ may associate with $a_1, p_0$, {\it etc.} If the user 
is not familiar with the implementation of this adder, those cross-literals will bring
confusions. However, if word-level techniques allow specification like Equation \ref{eqn:wordspec},
the verification tool will be very user-friendly and straightforward even if the implementation details
are in a black box.

\subsection{On the Existence of Word-Level Abstraction for Arbitrary Circuits}
When given a bit-level netlist, the prerequisites to use word-level techniques are to convert bit-level 
to word-level first. This conversion is usually completed by abstraction techniques.

An old but universally effective abstraction method is {\it Lagrange's interpolation}, 
which can be applied over finite fields. Here we use an example to illustrate the 
conversion in Galois field using Lagrange's interpolation for an arbitrary circuit.

@@

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$}
\end{table}
Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
$Z = \Func(A)$. Recall Lagrange's interpolation formula:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ dots with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

The Lagrange's interpolation theorem also proves the existence of a word-level abstraction for a bit-level netlist.
In practice, Lagrange's interpolation is not scalable. 
The reason is that it needs the entire function (state space), but usually we only have the circuit representation
of the FSM. Considering this fact, a symbolic method is needed.
Our approach in this section uses abstraction based on 
Gr\"obner basis with abstraction term order (ATO), which is briefly introduced in Section \ref{sec:abstraction}.

\subsection{Significance of Developing Word-Level Reachability Analysis}
% At the end of this section, we summarize each subsection and conclude the motivation of our research.
% First, we start from the basic FSM traversal algorithms and exhibit their
% difficulties to overcome. Secondly, we state the observation that word-level description is 
% increasingly important and common in characterizing the data flow on modern large-size datapaths. 
% Last but not least, we give instances that some prerequisites for supporting word-level techniques 
% are already available.

Based on the aforementioned discussions, the importance of performing FSM traversal at word-level 
can bring a dimension of abstraction in sequential circuit reachability analysis.
To overcome the cost incurred by searching in a large space, we propose to use word-level 
polynomials to represent the states and transition relations. As a result, states are categorized into 
sets represented by a small number of word-level polynomials (more specifically, the varieties to the polynomial ideals),
and multiple transition relations are therefore merged together. All of these efforts reduce the cost of 
state space, meanwhile lowering the time complexity to traverse such a state space. As Lagrange's interpolation
confirms the existence of word-level abstraction for bit-level circuits, the concept
can be extended to encode the state space of sequential circuit by word-level polynomials in $\Fkk$.
% we prove the feasibility of word-level techniques on general circuits' verification.
% Thus we cover both the necessity and sufficiency of developing such 
% a word-level traversal technique in our work.
% They are also answers to the questions at the beginning of this section.

@@

\subsection{FSM Model for Sequential Circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$
\begin{figure}[H]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{newfig/seqmodel.pdf}
% \vspace{-0.2in}
\caption{FSM models of sequential circuits}
%\end{minipage}
\label{fig:seqmodel}}
\end{figure}
In some cases, arithmetic computations are implemented as Moore machines where input operands
are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
We can simplify them to the model in Figure \ref{fig:seqmodel}(b).

@@

Here, $\exists x (f)$ represents the {\it existential quantification
  of $f$ {\it w.r.t.} variable $x$}. In Boolean logic, this operator is implemented as
  $$\exists x (f) = f_x\lor f_{\overline{x}}$$

@@

\subsection{FSM Traversal at Word-Level over $\Fkk$} 
The state transition graph (STG) shown in Figure \ref{fig:fsm}(a) uses a
2-bit Boolean vector to represent 4 states $\{S_0, S_1, S_2,
S_3\}$. We map these states to elements in $\mathbb{F}_{2^2}$, where
$S_0 = 0, S_1 = 1, S_2 = \alpha, S_3 = \alpha+1$. Here, we take 
$P(X) = X^2+X+1$ as the irreducible polynomial to construct
$\mathbb{F}_4$, and $P(\alpha) = 0$ so that $\alpha^2 + \alpha + 1 =
0$.

{\it Initial state:} Line 1 of Algorithm \ref{alg:BFS} specifies the initial state.
In algebraic geometry, it can be specified by means of a corresponding polynomial
 $f = \mathcal{F}(S) =
S - 1 - \alpha$. Notice that if we consider the ideal generated by the
initial state polynomial, $I = \langle f\rangle$, then its variety
$V(I) = 1+\alpha$ corresponds to the state encoding $S_3 = \{11\} =
1+\alpha$, where a polynomial in word-level variable $S$ encodes the
initial state. 
% with only one generator $f$, its variety
% $V(I) = \{\gamma\ |\ \gamma \in \mathbb{F}_{2^2}, \gamma = 1+\alpha\}$, which equals to $\{1+\alpha\}$, the only
% valid value $S_3$ can take.

@@

Let $x^{2^k}-x$ be a vanishing polynomial in $\Fkk[x]$. Then $V(x^{2^k}-x)=\Fkk$
{\it i.e.} the variety of vanishing ideal contains
 all possible valuations of variables, so it constitutes the {\bf
   universal set}. Subsequently, based on Theorem \ref{thm:quotient},
 the {\bf absolute complement} $V(J')$ of a variety $V(J)$ can be
 computed as:

@@

\section{Problem Setup and Formulation}
\begin{Problem}
We use the notions from Figure \ref{fig:seqmodel}(a) in this setup. 
\begin{enumerate}[{1)}] 
\item The circuit is modeled over $\Fkk = \F_2[x]\pmod{p(x)}, p(\alpha) = 0$, where $k$ is the number of flip-flops,
	the PS variables $\{s_0,dots,s_{k-1}\}$, NS variables $\{t_0,\dots,t_{k-1}\}$.
\item Denote $S$ as the PS word-level by the following polynomial:
$$f_S: S = s_0+s_1\alpha+s_2\alpha^2+\cdots+s_{k-1}\alpha^{k-1}$$
Similarly we define NS word-level variable $T$ by:
$$f_T: T = t_0+t_1\alpha+t_2\alpha^2+\cdots+t_{k-1}\alpha^{k-1}$$
\item Impose ATO for sequential circuits which is 
$$LEX: \text{ all bit-level variables in any order }>S>T$$
\item Write polynomials $f_1,\dots,f_s$ for each gate, and construct ideal describing the circuit:
$$J_{ckt} = \langle f_1,\dots,f_s,f_S,f_T\rangle$$
as well as the ideal with vanishing polynomials:
$$J_0 = \langle \dots,x_i^2-x_i,\dots,S^{2^k}-S,T^{2^k}-T\rangle$$
where $x_i$ corresponds to bit-level variables denoting wires in the circuit.
\item Compute GB with ATO for $G$. Then obtain the projection of variety only on NS variable $T$, by
$$G\cap \Fkk[T]$$
\end{enumerate}
\end{Problem}

@@

\subsection{Word-Level FSM Traversal Example}
\begin{Example}
\label{ex:SMPO}
We apply Algorithm \ref{alg:univa} to the example shown in
Figure \ref{fig:fsm} to execute the FSM traversal. Let the initial state
$from^0 = \{00\}$ in $\B^2$ or $0 \in \mathbb F_4$. Polynomially, it is
written as $from^0 = S - 0$. In the first iteration, we compose
an ideal $J$ with

@@

$V(\langle T^3 + 1\rangle) = \{1, \alpha, \alpha+1\}$ denoting the states
$\{01,10,11\}$. Then the newly reached state set in this iteration is 
$$\langle T^3+1, T^2+(\alpha+1)T+\alpha \rangle = \langle
T^2+(\alpha+1)T+\alpha \rangle$$ We add these states 
to formerly reached states 
\begin{align*}
reach &= \langle T\rangle \cdot \langle T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T\cdot T^2+(\alpha+1)T+\alpha \rangle \\
&= \langle T^3+(\alpha+1)T^2+\alpha T\rangle
\end{align*}
 {\it i.e.} states $\{00,01,10\}$. We update the present states
for next iteration $$from^1 = S^2+(\alpha+1)S+\alpha.$$
